---
title: "Homework 5"
author: "Eman Ibrahim (ei2291)"
output: github_document
    
---


```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(rvest)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

```{r}
set.seed(1)
```

Writing function to test
```{r}
birthday_sim = function(n){
 birthdays = sample(1:365, size = n, replace = TRUE)
 any(duplicated(birthdays))
}
```
test:
```{r}
birthday_sim(10)
```

Run the function 10,000 times for each group (2 to 50)
```{r}
sim_results = 
  expand_grid(
    group_size = 2:50,
    iter = 1:10000
  ) |> 
  mutate(
    shared = map_lgl(group_size, birthday_sim)
  )
```

Probability of shared birthdays:
```{r}
prob_results = 
  sim_results |> 
  group_by(group_size) |> 
  summarize(prob_shared = mean(shared))
```

Plot:
```{r}
prob_results |> 
  ggplot(aes(x=group_size, y=prob_shared))+
  geom_line(color="blue", linewidth = 1.2)+
  geom_point(size=1.2)+
  labs(
    title = "Probability of shared birthdays by group size",
    x= "Group size",
    y= "Probability"
  )+
  theme_minimal()
```

Based on the result of our plot, we can see that the probability of a shared birthday increases as group size increases. At about a group size of 23, the chance(probability) is about 50%. By around a group size of 50, there is almost 100% chance of having a shared birthday. We see that the probability increases very fast even when there are many possible birthdays. 





